Supporting information: Copulas and their potential for ecology

### 2019-10-06 Paper.Rmd
working on Q3 edits

<!--Second mechanism, the Moran one-->
The hypothesis that characteristics of 
the copula structure of spatial dependence 
in an ecological variable can be inherited from an environmental 
variable was found to be reasonable, because it held for the models we examined; 
however, similarities between environmental- and 
ecological-variable copula structure were reduced when 
dynamics were strongly nonlinear.
For simulations using AR(1) models, our correlation and tail asymmetry statistics 
were always similar for noise and populations 
(Fig. \ref{fig_Cause4copula_C_kend_ar1} for the Clayon copula, 
Fig. \ref{SM-fig_Cause4copula_SC_kend_ar1} for the survival Clayton copula).
<!--***DAN: reorder sup mat figs so they are cited in order-->
Though there were significant differences for many statistics
and simulations, these were small compared to the overall tendency for
larger (respectively, smaller) values of our statistics, as applied to
noise, to be paired with larger (respectively, smaller) values of the same
statistics as applied to model outputs. 
For our nonlinear model with weak noise, values of the statistics
were again quite similar for noise and model outputs 
(Fig. \ref{SM-fig_Cause4copula_C_kend_sRicker_nsd_0.1} for Clayton, 
Fig. \ref{SM-fig_Cause4copula_SC_kend_sRicker_nsd_0.1} for survival Clayton).
Though there were again significant differences for many statistics
and simulations, these were again small relative to overall variation of 
values of the statistics.
Since many ecological models are nonlinear, this result provides the 
reasonable expectation that there may be a Moran-effect-like correspondence 
between noise and model-output dependence structure across space 
for typical ecological dynamics,
as long as environmental noise is small enough that dynamics stay relatively close
to the model equilibrium. Results that hold for "weak noise" in this sense
are common in ecology.
For strong noise and using our nonlinear model, 
our correlation and tail-asymmetry statistics, generally 
speaking, were approximately similar between noise and model
outputs; however, similarity was reduced relative to previous results, 
and for a few simulations, asymmetry statistics
had opposite signs for noise and model outputs 
(Figs \ref{SM-fig_Cause4copula_C_kend_sRicker_nsd_1} and \ref{SM-fig_Cause4copula_SC_kend_sRicker_nsd_1}). For instance, 
using a Clayton  copula with a large Kendall $\tau$,
$\cor_l - \cor_u$ was slightly positive for noise, but slightly negative for 
population model outputs (Fig. \ref{SM-fig_Cause4copula_C_kend_sRicker_nsd_1}).

@@

The hypothesis was presented in the Introduction that
the distribution (through time) of a spatially 
averaged quantity should be influenced by 
dependencies between the local quantities being averaged, including their
copula structure and tail dependence. 
To refine this hypothesis, suppose an ecological variable 
$E_i(t)$ is measured at locations $i=1,\ldots,N$ and times $t=1,\ldots,T$, and
the spatial mean $\sum_i E_i(t)/N$ is of
interest. The $E_i(t)$ could be, for instance, local abundances of a pest or 
exploited species, or local fluxes of a greenhouse gas. If  $E_i$ and $E_j$ are 
right-tail dependent for most location pairs $i$ and $j$, then
exceptionally large values tend to occur at the same time in most locations.
We hypothesize that this 
will tend to increase the skewness of the distribution 
of the spatial mean. Similarly, left-tail dependence between local variables
should tend to decrease skewness. Strong positive skewness of the 
spatial-mean time series corresponds to "spikiness" of that time series,
which corresponds to instability through time. 
The spatial-mean time series and its skewness may be quantities of 
principal importance for 
pest or resource abundance, for which extreme values (spikes) 
in the spatial mean have large effects.

To help answer Q3, we tested the above hypothesis
using our multivariate datasets (Table \ref{tab_data_info}).
For each dataset we calculated the spatial mean time series, and then 
the skewness through time of that
mean. Then, for each dataset, we compared the value obtained 
to a distribution of values 
of the same quantity for each of 10000 surrogate datasets. Surrogate 
datasets were produced by randomizing the empirical data
in a special way to have the 
copula structure of a multivariate-normal distribution, but 
to retain exactly the same
distributions of values for each sampling location as the original data
(Appendix \ref{SM-Surrog_multi}). 
Surrogates also had very similar Spearman correlations 
between pairs of sampling locations
as the data. 
Our comparisons therefore tested the null 
hypothesis that the skewness of the spatial 
mean took values on the empirical data no different 
from what one would expect if the copula
structure of the data were multivariate N, but the data were otherwise statistically
unchanged. Significant differences indicate that non-normal copula structure in the data 
contributed to the skewness of the spatial mean time series, 
i.e., to its instability and "spikiness" through time.

For green spruce aphid abundance data, *C. furca* abundance data, and methane-flux
data, because these datasets exhibited lower-tail dependence 
(Table \ref{tab_multivar_summary}), we compared empirical and surrogate skewness values
via a one-tailed test in the left tail: the $p$-value was the fraction
of surrogate skewnesses less than the skewness for the empirical data; the test
examines whether lower-tail dependence caused the spatial average to have significantly lower
skewness than would have been expected without tail dependence. For
leaf-curling plum aphid data, because that dataset exhibited upper-tail 
dependence (Table \ref{tab_multivar_summary}), we did the analogous one-tailed test
in the right tail. The test examines whether upper-tail dependence caused the spatial
average to have significantly higher skewness than 
expected without tail dependence.

To further address Q3, we also examined a hypothesis that spatial 
tail dependence of an environmental variable can influence the 
extinction risk of a metapopulation. We hypothesized that environmental noises
exhibiting greater left-tail dependence across different habitat patches would 
cause higher metapopulation extinction risks because then very bad years for the 
component populations would occur simultaneously in many patches,
reducing rescue effects. Here we assume, for simplicity, 
that low values of the environmental
variable are "bad" for the populations and high values are "good".
We tested the reasonableness of this hypothesis using a metapopulation
extension of the Lewontin-Cohen model,
$\vec{P}(t+1) = D \lambda(t) \vec{P}(t)$,
where the $i^{th}$ component of the length-$N$ vector 
$\vec{P}(t)$ represents
population density in the $i^{th}$ habitat patch at time $t$. 
The $N \times N$ matrix $\lambda(t)$ was diagonal with $i^{th}$ diagonal entry 
$\exp(r+\epsilon_i(t))$. Here $r$ is a growth rate; we used $r=0$. The
$\epsilon_i(t)$ represent environmental noises. They were standard-normally 
distributed, were independent through time, and showed the same 
spatial correlations for every simulation, 
but were made to exhibit right- or left-tail dependence between patches in
different simulations (Appendix \ref{SM-ext_risk_model}). The $N \times N$
matrix $D$ was a dispersal matrix
modelling local or global dispersal at rate $d$, in different simulations
(Appendix \ref{SM-ext_risk_model}).
After each step, 
if the density in a patch was $<1$, it was
set to $0$. We simulated the model 10000 times for each combination of parameters,
starting from $p_0 = 50$ in each patch, 
and calculated extinction risk after $25$ time steps.

@@

\noindent Results were consistent with the hypothesis that the skewness
of a spatial-average time series is influenced by tail dependence
between the local quantities being averaged. Thus copula 
structure and tail dependence are important for spatially averaged 
quantities. This result is represented in Fig. \ref{MasterFigure}
as the solid box around "Instability/skewness of mean or total 
time series" and the solid arrows labeled "X". For datasets that
exhibited strong lower-tail (respectively, upper-tail) 
dependence in earlier analyses (green spruce 
aphid abundance data for lower-tail, and leaf-curling plum aphid first flight data
for upper-tail, Table 
\ref{tab_multivar_summary}), skewness of the spatial average 
was less than (respectively, greater than) a significant fraction of 
surrogate skewness values 
(Fig. \ref{fig_skewness_spearman}A,B). For datasets with moderate lower-tail 
dependence (\emph{Ceratium furca} abundance and methane data), 
skewness of the spatial
average showed a non-significant or marginally significant tendency 
toward being less than surrogate skewnesses (Fig. \ref{fig_skewness_spearman}C,D).

\begin{figure}[!h]
\textbf{ \hspace{6.6 cm} (A) \hspace{4.6 cm} (B)} \\
\vspace{-0.4 cm}
\begin{center}
\includegraphics[width=5 cm]{./Results/skewness_results/skewness_aphid_count/skewness_aphid_count_sp_10_spearman.pdf}
\includegraphics[width=5 cm]{./Results/skewness_results/skewness_aphid_ff/skewness_aphid_ff_sp_11_spearman.pdf}\\
\end{center}
\vspace{0.2 cm}
\textbf{ \hspace{6.6 cm} (C) \hspace{4.6 cm} (D)} \\
\vspace{-0.4 cm}
\begin{center}
\includegraphics[width=5 cm]{./Results/skewness_results/skewness_plankton_north_sea/skewness_plankton_north_sea_sp_16_spearman.pdf}
\includegraphics[width=5 cm]{./Results/skewness_results/skewness_methane/skewness_methane_spearman.pdf}
\caption{Skewness of spatially averaged green spruce aphid abundance $(A)$, leaf-curling plum 
aphid first flight dates $(B)$, \emph{Ceratium furca} abundance $(C)$, and methane-flux
$(D)$ time series compared to a multivariate 
normal-copula null hypothesis. Black dots are empirical 
skewnesses; see text 
and Appendix \ref{SM-Surrog_multi} for details of the null hypothesis. Results show a 
tendency for skewness of the spatial average to be affected as hypothesized 
by tail dependence. \label{fig_skewness_spearman}}
\end{center}
\end{figure}

Consistent with our extinction risk hypothesis, 
left-tail-dependent environmental fluctuations increased metapopulation
extinction risk for the spatial 
Lewontin-Cohen model (Fig. \ref{fig_extrisk_lcohen_local}), 
for $N=5$ and $N=25$ habitat patches, and for local 
(Fig. \ref{fig_extrisk_lcohen_local}) and global (Fig. \ref{SM-fig_extrisk_lcohen_global})
dispersal. This result is represented in Fig. \ref{MasterFigure}
as the dashed box around "Extinction risk" and the dashed arrow labelled "Y".
We also substantiated using models (Appendix \ref{SM-Consequence_TL_simulation}, \ref{SM-sect:copsurrognd}, Fig. \ref{SM-fig_spTL})
a hypothesis (Fig. \ref{MasterFigure}, arrow labelled "Z") that 
Taylor's law can be influenced by
tail dependence.

\begin{figure}[!h]
\textbf{ \hspace{6.6 cm} (A) \hspace{4.2 cm} (B)} \\
\vspace{-0.8 cm}
\begin{center}
\includegraphics[width=10 cm]{./Results/ext_risk_copula/lcohen_ext_risk_local_disp.pdf}
\caption{Extinction risk for the metapopulation extension of the Lewontin-Cohen model, after 
$25$ time steps, was higher for left-tail-dependent environmental noise. Dispersal was local,
with dispersal rate $d$ (Appendix \ref{SM-ext_risk_model}). Simulations used $N$ patches
for $N=5$ (A) and $N=25$ (B).\label{fig_extrisk_lcohen_local}}
\end{center}
\end{figure}

>>>>

<!--Second mechanism, the Moran one-->
The hypothesis that characteristics of 
the copula structure of spatial dependence 
in an ecological variable can be inherited from an environmental 
variable was found to be reasonable, because it held for the models we examined; 
however, similarities between environmental- and 
ecological-variable copula structure were reduced when 
dynamics were strongly nonlinear.
For simulations using AR(1) models, our correlation and tail asymmetry statistics 
were always similar for noise and populations 
(Fig. \ref{fig_Cause4copula_C_kend_ar1} for the Clayon copula, 
Fig. \ref{SM-fig_Cause4copula_SC_kend_ar1} for the survival Clayton copula).
Though there were significant differences for many statistics
and simulations, these were small compared to the overall tendency for
larger (respectively, smaller) values of our statistics, as applied to
noise, to be paired with larger (respectively, smaller) values of the same
statistics as applied to model outputs. 
For our nonlinear model with weak noise, values of the statistics
were again quite similar for noise and model outputs 
(Fig. \ref{SM-fig_Cause4copula_C_kend_sRicker_nsd_0.1} for Clayton, 
Fig. \ref{SM-fig_Cause4copula_SC_kend_sRicker_nsd_0.1} for survival Clayton).
Though there were again significant differences for many statistics
and simulations, these were again small relative to overall variation of 
values of the statistics.
Since many ecological models are nonlinear, this result provides the 
reasonable expectation that there may be a Moran-effect-like correspondence 
between noise and model-output dependence structure across space 
for typical ecological dynamics,
as long as environmental noise is small enough that dynamics stay relatively close
to the model equilibrium. Results that hold for "weak noise" in this sense
are common in ecology.
For strong noise and using our nonlinear model, 
our correlation and tail-asymmetry statistics, generally 
speaking, were approximately similar between noise and model
outputs; however, similarity was reduced relative to previous results, 
and for a few simulations, asymmetry statistics
had opposite signs for noise and model outputs 
(Figs \ref{SM-fig_Cause4copula_C_kend_sRicker_nsd_1} and \ref{SM-fig_Cause4copula_SC_kend_sRicker_nsd_1}). For instance, 
using a Clayton  copula with a large Kendall $\tau$,
$\cor_l - \cor_u$ was slightly positive for noise, but slightly negative for 
population model outputs (Fig. \ref{SM-fig_Cause4copula_C_kend_sRicker_nsd_1}).

@@

The hypothesis was presented in the Introduction that
the distribution (through time) of a spatially 
averaged quantity should be influenced by 
dependencies between the local quantities being averaged, including their
copula structure and tail associations. 
To refine this hypothesis, suppose an ecological variable 
$E_i(t)$ is measured at locations $i=1,\ldots,N$ and times $t=1,\ldots,T$, and
the spatial mean $\sum_i E_i(t)/N$ is of
interest. The $E_i(t)$ could be, for instance, local abundances of a pest or 
exploited species, or local fluxes of a greenhouse gas. If  $E_i$ and $E_j$ are 
associated primarily in their right tails for most location pairs $i$ and $j$, then
exceptionally large values tend to occur at the same time in most locations.
We hypothesize that this 
will tend to increase the skewness of the distribution 
of the spatial mean. Similarly, left-tail associations between local variables
should tend to decrease skewness. Strong positive skewness of the 
spatial-mean time series corresponds to "spikiness" of that time series,
which corresponds to instability through time. 
The spatial-mean time series and its skewness may be quantities of 
principal importance for 
pest or resource abundance, for which extreme values (spikes) 
in the spatial mean may have large effects.

To help answer Q3, we tested the above hypothesis
using our multivariate datasets (Table \ref{tab_data_info}).
For each dataset we calculated the spatial mean time series, and then 
the skewness through time of that
mean. Then, for each dataset, we compared the value obtained 
to a distribution of values 
of the same quantity for each of 10000 surrogate datasets. Surrogate 
datasets were produced by randomizing the empirical data
in a special way to have the 
copula structure of a multivariate-normal distribution, but 
to retain exactly the same
distributions of values for each sampling location as the original data
(Appendix \ref{SM-Surrog_multi}). 
Surrogates also had very similar Spearman correlations 
between pairs of sampling locations
as the data. 
Our comparisons therefore tested the null 
hypothesis that the skewness of the spatial 
mean took values on the empirical data no different 
from what one would expect if the copula
structure of the data were multivariate normal (i.e., the same copula as a 
multivariate normal distribtion), but the data were otherwise statistically
unchanged. Significant differences indicate that non-normal copula structure in the data 
contributed to the skewness of the spatial mean time series, 
i.e., to its instability and "spikiness" through time.

For green spruce aphid abundance data, *C. furca* abundance data, and methane-flux
data, because these datasets exhibited stronger lower- than upper-tail associations 
(Table \ref{tab_multivar_summary}), we compared empirical and surrogate skewness values
via a one-tailed test in the left tail: the $p$-value was the fraction
of surrogate skewnesses less than the skewness for the empirical data. The test
examines whether stronger lower-tail associations between local time series 
caused the spatial average to have significantly lower
skewness than would have been expected with symmetric tail associations. For
leaf-curling plum aphid data, because that dataset exhibited stronger upper-
than lower-tail 
associations (Table \ref{tab_multivar_summary}), we did the analogous one-tailed test
in the right tail. The test examines whether upper-tail associations caused the spatial
average to have significantly higher skewness than 
would have been expected with symmetric tail associations.

To further address Q3, we also examined a hypothesis that 
asymmetric tail associations across space of an environmental variable can influence the 
extinction risk of a metapopulation. We hypothesized that environmental noises
exhibiting greater left-tail associations across different habitat patches would 
cause higher metapopulation extinction risks because then very bad years for the 
component populations would occur simultaneously in many patches,
reducing rescue effects. Here we assume, for simplicity, 
that low values of the environmental
variable are "bad" for the populations and high values are "good".
We tested the reasonableness of this hypothesis using a metapopulation
extension of the Lewontin-Cohen model,
$\vec{P}(t+1) = D \lambda(t) \vec{P}(t)$,
where the $i^{th}$ component of the length-$N$ vector 
$\vec{P}(t)$ represents
population density in the $i^{th}$ habitat patch at time $t$. 
The $N \times N$ matrix $\lambda(t)$ was diagonal with $i^{th}$ diagonal entry 
$\exp(r+\epsilon_i(t))$. Here $r$ is a growth rate; we used $r=0$. The
$\epsilon_i(t)$ represent environmental noises. They were standard-normally 
distributed, were independent through time, and showed the same 
spatial correlations for every simulation, 
but were made to exhibit stronger right- or left-tail associations between patches in
different simulations (Appendix \ref{SM-ext_risk_model}). The $N \times N$
matrix $D$ was a dispersal matrix
modelling local or global dispersal at rate $d$, in different simulations
(Appendix \ref{SM-ext_risk_model}).
After each step, 
if the density in a patch was $<1$, it was
set to $0$. We simulated the model 10000 times for each combination of parameters,
starting from $p_0 = 50$ in each patch, 
and calculated extinction risk after $25$ time steps.

<!--***DAN: if you do decide to move the TL stuff to main text, this is where the 
methods go. Other than this, I think I am done revising methods for Q3.-->

<!--***DAN: Just ike in the Causes secton you added a comment saying these models
were for initial exploration only and so we did not explore parameter space
nor look at model variants, likewise we also need some commentary in this section 
indicating that the modelling is simple and not comprehensively explored, because
the purpose is just to get an idea of what kinds of consequences of copula 
structure *may* pertain; further work is needed to understand to what extent these
consequences are realized in real systems.-->

Pursuing ideas that sprang from earlier theoretical work [@Reuman2017], we 
tested whether the copula structure of the dependences between population variables 
measured in different locations influences the spatial version of Taylor's
law. Taylor's law is a commonly observed and widely studied 
[@Taylor1961; @Taylor1988; @Cohen2013; @Xu2016; @Reuman2017] 
empirical pattern that relates the 
variances of groups, $g$, of population measurements to the means of the 
groups via a power law, $v_g = a \times m_g^b$, or equivalently, 
$\log(v_g)=\log(a)+b \times \log(m_g)$. Here $b$ is called 
the Taylor's law *exponent* or 
*slope*, and $\log(a)$ is the Taylor's law intercept. 
There are several versions of Taylor's law. For spatial Taylor's law,
given population density or abundance data $x_i(t)$ measured in locations
$i=1,\ldots,N$ and times $t=1,\ldots,T$, a group $g$ consists of 
all the measurements $x_i(t)$, $i=1,\ldots,N$ made in different locations at the same time.
So means $m_g$ and variances $v_g$ are computed across space. We consider that 
Taylor's law is validated for a dataset if the 
$\log(v_g)$ versus $\log(m_g)$ scatter plot 
for the dataset shows a linear, homoskedastic pattern.
Taylor's law has been 
verified empirically for hundreds of taxa and has been applied in numerous fields
including fisheries management, estimation of species persistence times, and
agriculture [@Cohen2015; @Reuman2017].

@@

\noindent Results were consistent with the hypothesis that the skewness
of a spatial-average time series is influenced by tail associations
between the local quantities being averaged. 
For datasets that
exhibited highly asymmetric tail associations in earlier analyses (green spruce 
aphid abundance data had stronger lower-tail association, and, 
respectively,  leaf-curling plum 
aphid first flight data had stronger upper-tail association, Table 
\ref{tab_multivar_summary}), skewness of the spatial average 
was less than (respectively, greater than) a significant fraction of 
surrogate skewness values 
(Fig. \ref{fig_skewness_spearman}A,B). For datasets with moderately stronger 
lower than upper-tail 
dependence (\emph{Ceratium furca} abundance and methane data), 
skewness of the spatial
average showed a non-significant or marginally significant tendency 
toward being less than surrogate skewnesses (Fig. \ref{fig_skewness_spearman}C,D). 
Thus copula 
structure and asymmetric tail associations are important for spatially averaged 
quantities. This result is represented in Fig. \ref{MasterFigure}
as the solid box around "Instability/skewness of mean or total 
time series" and the solid arrows labeled "X".

\begin{figure}[!h]
\textbf{ \hspace{6.6 cm} (A) \hspace{4.6 cm} (B)} \\
\vspace{-0.4 cm}
\begin{center}
\includegraphics[width=5 cm]{./Results/skewness_results/skewness_aphid_count/skewness_aphid_count_sp_10_spearman.pdf}
\includegraphics[width=5 cm]{./Results/skewness_results/skewness_aphid_ff/skewness_aphid_ff_sp_11_spearman.pdf}\\
\end{center}
\vspace{0.2 cm}
\textbf{ \hspace{6.6 cm} (C) \hspace{4.6 cm} (D)} \\
\vspace{-0.4 cm}
\begin{center}
\includegraphics[width=5 cm]{./Results/skewness_results/skewness_plankton_north_sea/skewness_plankton_north_sea_sp_16_spearman.pdf}
\includegraphics[width=5 cm]{./Results/skewness_results/skewness_methane/skewness_methane_spearman.pdf}
\caption{Skewness of spatially averaged green spruce aphid abundance $(A)$, leaf-curling plum 
aphid first flight dates $(B)$, \emph{Ceratium furca} abundance $(C)$, and methane-flux
$(D)$ time series compared to a multivariate 
normal-copula null hypothesis. Black dots are empirical 
skewnesses; see text 
and Appendix \ref{SM-Surrog_multi} for details of the null hypothesis. Results show a 
tendency for skewness of the spatial average to be affected as hypothesized 
by asymmetric tail associations. \label{fig_skewness_spearman}}
\end{center}
\end{figure}

Consistent with our extinction risk hypothesis, 
left-tail-associated environmental fluctuations increased metapopulation
extinction risk for the spatial 
Lewontin-Cohen model, 
for $N=5$ and $N=25$ habitat patches, and for local 
and global
dispersal (Fig. \ref{fig_extrisk_lcohen_local}). This result 
is represented in Fig. \ref{MasterFigure}
as the dashed box around "Extinction risk" and the dashed arrow labelled "Y".

<!--***DAN: the below is on TL and needs to be augmented if we keep TL results-->
We also substantiated using models (Appendix \ref{SM-Consequence_TL_simulation}, \ref{SM-sect:copsurrognd}, Fig. \ref{SM-fig_spTL})
a hypothesis (Fig. \ref{MasterFigure}, arrow labelled "Z") that 
Taylor's law can be influenced by
tail dependence.

\begin{figure}[!h]
\textbf{ \hspace{3.6 cm} (A) \hspace{4.2 cm} (B)} \\
\vspace{-0.8 cm}
\begin{center}
\includegraphics[width=10 cm]{./Results/ext_risk_copula/lcohen_ext_risk_local_disp.pdf}
\end{center}
\textbf{ \hspace{3.6 cm} (C) \hspace{4.2 cm} (D)} \\
\vspace{-0.8 cm}
\begin{center}
\includegraphics[width=10 cm]{./Results/ext_risk_copula/lcohen_ext_risk_global_disp.pdf}
\end{center}
\caption{Extinction risk for the metapopulation extension of the Lewontin-Cohen model, after 
$25$ time steps, was higher for environmental noise with stronger left-tail association
across space. Dispersal was local (A, B) or global (C, D),
with dispersal rate $d$ (Appendix \ref{SM-ext_risk_model}). Simulations used $N$ patches
for $N=5$ (A, C) and $N=25$ (B, D).\label{fig_extrisk_lcohen_local}}
\end{figure}

### 2019-10-06 SupportingInformation.Rmd
working on Q3 edits

<!--***Methods for generating evolutionary changes, keep this here-->
\noindent The matrices from which root character states and changes 
across phylogeny branches were randomly chosen were generated as follows.
For extreme right-tail dependent noise, we first produced
a 1,000,000 by $2$ matrix, $m$, filled with 2,000,000 
independent standard-normal random draws. Then, each row of $m$ 
was identified for a first set of modifications randomly 
and independently with a $50\%$ chance; 
for those rows identified, every element of the row was replaced with the 
absolute value of the first element of the row (including the first element itself).
For other rows, every element in the row was replaced by the negative of its
own absolute value. This produces right-tail dependent noise, normally distributed
for each of the two locations. 
If left-tail dependent noise was desired, we took the negative of the
entire matrix. The moderately left-tail dependent matrix was based on a Clayton copula 
with parameter chosen (using the `iRho` function in the `copula` package) so that
the Spearman correlation would be the same as for the extreme tail dependence 
datasets. The moderately right-tail dependent matrix and the symmetric tail dependence
matrix were constructed similarly, but using an survival Clayton copula and an normal copula,
respectively.

<!--Move all the rest of it into the Discussion, I think, but can decide when I get to the Discussion. If it is mved, then you may want to retitle this appendix to reflect the 
fact it is onl methods-->
Performing statistical tests for associations between continuous traits in different species
    was a primary motivating example of phylogenetic comparative methods.
For example, Felsenstein's method of phylogenetically independent contrasts 
    [@Felsenstein1985] is currently the second most-cited paper in the history
    of the journal the American Naturalist [@HueyGT2019].
Frustratingly, the field still lacks a dependable procedure for dealing with branch lengths, 
    which are a crucial input to the method.
Ideally, the branch lengths used to correct for phylogenetic effects would represent
    the expected amounts of change for the characters that are being analyzed.
Because researchers almost never have a reliable method for providing such branch lengths,
    most researchers rely on ultrametric trees -- those for which the branch
    length can be treated as a proxy for the duration of the branch in time.
Frequently these branch lengths are transformed to assess sensitivity to different 
    assumptions about the degree of phylogenetic inertia displayed by the traits
    under study [@Ives2018;@Harmon2018].
Even if one were able to simply use a time-based set of branch lengths, assigning
    dates to nodes in phylogenies is difficult.
DNA sequence data can provide estimates of branch lengths, but these estimates
    are dependent on the adequacy of models which correct sequences for multiple
    substitutions occurring at the same location.
Biases in estimating the evolutionary distance can affect downstream analyses [@Phillips2009].
Additionally, changes
    in the rate of molecular evolution make the estimation of dates difficult [@HeathM2014] even when branch lengths in time are accurately estimated.

@@

<!--adding plots showing effects of asymmetric copula on extinction risk-->
\begin{figure}[!h]
\textbf{ \hspace{7.3 cm} (A) \hspace{4 cm} (B)} \\
\vspace{-0.8 cm}
\begin{center}
\includegraphics[width=10 cm]{./Results/ext_risk_copula/lcohen_ext_risk_global_disp.pdf}
\caption[Analogous to Fig. \ref{MT-fig_extrisk_lcohen_local} in the main text, but with global dispersal]{Analogous to Fig. \ref{MT-fig_extrisk_lcohen_local} in the main text, but with global dispersal.\label{fig_extrisk_lcohen_global}}
\end{center}
\end{figure}

@@

\noindent Pursuing ideas that sprang from earlier theoretical work [@Reuman2017], we 
tested whether the copula structure of the dependences between population variables 
measured in different locations influences the spatial version of Taylor's
law. Taylor's law is an extremely commonly observed and widely studied 
[@Taylor1961; @Taylor1988; @Cohen2013; @Xu2016] empirical pattern that relates the 
variances of groups, $g$, of population measurements to the means of the 
groups via a power law, $v_g = a \times m_g^b$, or equivalently, 
$\log(v_g)=\log(a)+b \times \log(m_g)$. Here $b$ is called 
the Taylor's law *exponent* or 
*slope*, and we refer to $\log(a)$ as the Taylor's law intercept. 
There are several versions of Taylor's law. For spatial Taylor's law,
given population density or abundance data $x_i(t)$ measured in locations
$i=1,\ldots,N$ and times $t=1,\ldots,T$, a group $g$ consists of 
all the measurements $x_i(t)$, $i=1,\ldots,N$ made in different locations at the same time.
So means $m_g$ and variances $v_g$ are computed across space. We view the assumptions
of Taylor's law to be validated for a dataset if the 
$\log(v_g)$ versus $\log(m_g)$ scatter plot 
for the dataset shows a linear, homoskedastic pattern.
Taylor's law has been 
verified empirically for hundreds of taxa and has been applied in numerous fields
including fisheries management, estimation of species persistence times, and
agriculture [@Cohen2015; @Reuman2017].

>>>>

\noindent The matrices from which root character states and changes 
across phylogeny branches were randomly chosen were generated as follows.
For extreme right-tail dependent noise, we first produced
a 1,000,000 by $2$ matrix, $m$, filled with 2,000,000 
independent standard-normal random draws. Then, each row of $m$ 
was identified for a first set of modifications randomly 
and independently with a $50\%$ chance; 
for those rows identified, every element of the row was replaced with the 
absolute value of the first element of the row (including the first element itself).
For other rows, every element in the row was replaced by the negative of its
own absolute value. This produces right-tail dependent noise, normally distributed
for each of the two locations. 
If left-tail dependent noise was desired, we took the negative of the
entire matrix. The moderately left-tail dependent matrix was based on a Clayton copula 
with parameter chosen (using the `iRho` function in the `copula` package) so that
the Spearman correlation would be the same as for the extreme tail dependence 
datasets. The moderately right-tail dependent matrix and the symmetric tail dependence
matrix were constructed similarly, but using an survival Clayton copula and an normal copula,
respectively.

<!--***DAN: Move all the rest of it into the Discussion, I think, but can decide when I get to the Discussion. If it is mved, then you may want to retitle this appendix to reflect the 
fact it is onl methods-->
Performing statistical tests for associations between continuous traits in different species
    was a primary motivating example of phylogenetic comparative methods.
For example, Felsenstein's method of phylogenetically independent contrasts 
    [@Felsenstein1985] is currently the second most-cited paper in the history
    of the journal the American Naturalist [@HueyGT2019].
Frustratingly, the field still lacks a dependable procedure for dealing with branch lengths, 
    which are a crucial input to the method.
Ideally, the branch lengths used to correct for phylogenetic effects would represent
    the expected amounts of change for the characters that are being analyzed.
Because researchers almost never have a reliable method for providing such branch lengths,
    most researchers rely on ultrametric trees -- those for which the branch
    length can be treated as a proxy for the duration of the branch in time.
Frequently these branch lengths are transformed to assess sensitivity to different 
    assumptions about the degree of phylogenetic inertia displayed by the traits
    under study [@Ives2018;@Harmon2018].
Even if one were able to simply use a time-based set of branch lengths, assigning
    dates to nodes in phylogenies is difficult.
DNA sequence data can provide estimates of branch lengths, but these estimates
    are dependent on the adequacy of models which correct sequences for multiple
    substitutions occurring at the same location.
Biases in estimating the evolutionary distance can affect downstream analyses [@Phillips2009].
Additionally, changes
    in the rate of molecular evolution make the estimation of dates difficult [@HeathM2014] even when branch lengths in time are accurately estimated.

@@

<!--adding plots showing effects of asymmetric copula on extinction risk-->
<!--Combined this with the main text fig for AER
\begin{figure}[!h]
\textbf{ \hspace{7.3 cm} (A) \hspace{4 cm} (B)} \\
\vspace{-0.8 cm}
\begin{center}
\includegraphics[width=10 cm]{./Results/ext_risk_copula/lcohen_ext_risk_global_disp.pdf}
\caption[Analogous to Fig. \ref{MT-fig_extrisk_lcohen_local} in the main text, but with global dispersal]{Analogous to Fig. \ref{MT-fig_extrisk_lcohen_local} in the main text, but with global dispersal.\label{fig_extrisk_lcohen_global}}
\end{center}
\end{figure}
-->

@@
